\chapter{Orders of operations}
\label{chap:ooo}
A common question is ``ought I perform X before Y?''
The vigilant reader ought by now be able to give the answers.

\section{Transforming Pokémon}
\label{sec:pokemontransforms}
To minimize cost of Pokémon development, proceed in this order:
\begin{itemize}
\item Purify (saves costs, necessary for trade)
\item Trade (might get Lucky trade, saving costs)
\item Teach second attack (evolution might add cost)
\item Evolve and power up interchangeably
\end{itemize}

\section{Prioritizing encounters}
\label{sec:prioritymap}
The Trainer must often select between numerous possible interactions available from the map.
Especially when actively traveling, they must be prioritized, with some potentially missed.
I operate thusly:
\begin{itemize}
\item Do anything I particularly feel like at the moment.
      Ignore anything that irritates me\footnote{I expect never again to actively encounter Komala.}.
      This is a game, not a job.
\item Get anything I want that will otherwise be scrolled off the map.
\item Capture anything I actively want for battle/development/Research.
\item Spin a Pokéstop if I haven't since my last encounter (necessary to sustain ball inventory),
       or know myself to be short on balls.
\item Capture any Pokémon awarding more than minimum Stardust.
\item Capture anything missing from my Pokédex.
\item Capture any Pokémon awarding more than minimum Candy.
\item Capture anything boosted by my Mega/Primal form.
\item Capture anything weather-boosted.
\item Capture anything with a Max form I might use.
\end{itemize}
Below this level, if the Pokémon escapes once or twice, and I'm on the move,
I probably let it go.
\begin{itemize}
\item Capture anything I might want for battle/development.
\item Spin Pokestops (unless my bag is full).
\item Fight Team GO Rocket, if I feel like it.
\item Capture anything for which I lack platinum capture medal.
\item Capture anything known to be easy to catch.
\item Fight Max Battles or raids, if I feel like it.
\item Assault gyms, if I feel like it.
\item Feed berries to gym defenders for Stardust.
\item Continue idly depopulating the countryside.
\item Send gifts in celebration.
\end{itemize}
Throughout, I'm actively tossing as many berries as possible, and
 transferring Pokémon as rapidly as they come in.

\section{Prioritizing development}
\label{sec:developmentpriority}
What ought be evolved, what ought be powered up, and what ought be transferred?
A frontline is needed for raiding, Max Battles, Master League, and keeping Team Go Rocket out of one's face.
For each attack type, this ought eventually contain at least six (and ideally twelve) Pokémon
  employing powerful attacks of that type.
A Pokémon with a high level is much more valuable in the short term than one with a higher IV.
Until a minimal frontline is built up, don't worry about IVs; focus instead on strong species,
  good attacks, and levels sufficient to beat Team GO Rocket.
Once you're easily soloing first tier raids and Max Battles, begin collecting for the long term.
Transfer most of what you capture, keeping only quality species with decent IVs
  and captures of level sufficient to immediately serve on your frontline.
As you develop quality Pokémon, give your veterans the thanks they deserve: a quick trip to the Professor!
Savor the Candy through which they serve you even in death.

Trainers interested in League play will want a similar (but smaller) set of Pokémon chosen for each League's CP limits.
Read on\ldots
