\chapter{Optimal $\mathbfit{C\kern -1pt P}$-bound configurations\label{chap:optimal}}
\section[\CP1500 bound sorted by $\frac{\mathit{BS}}{3}$ (Great League)]%
{$\mathbfit{C\kern -1ptP}$\,1500 bound sorted by $\frac{\mathbfit{BS}}{\mathbf3}$ (Great League)\label{sec:cp1500a}}

Forms\footnote{Forms disallowed in League play are not listed.} are sorted by $\frac{\mathit{BS}}{3}$,
  the arithmetic mean of \Eff{A}, \Eff{D}, and \MHP\@.
This metric usually distinguishes between standard and Shadow forms (the geometric mean is always unchanged).
I include $\sqrt[3]{\BP\,}$ and even \CP\@.
These rankings \textit{do not} imply that a higher $\sqrt[3]{\BP\,}$ or $\frac{\mathit{BS}}{3}$
  will defeat a lower one.
Other issues---type effectiveness, attack choice, strategy---are usually more important.
Multiple sets of IVs and levels can score equally due to the discrete nature of \MHP\@.
In this case, all optimal sets are listed.

$A\%$ is the optimal configuration's advantage in terms of $\frac{BS}{3}$
  over the pessimal ``reasonable'' configuration, i.e. some IV
  and the highest level attainable given the \CP{} bound.
For \CP1500 (\autoref{table:cp1500a}, \autoref{table:cp1500g}), when a form can max out, i.e.\ is not restricted by the \CP{} bound,
  the maximum advantage tends to be 10\%--25\%, with a few outliers around 50\%.
It otherwise quickly drops to 4\% or less.
In \CP2500 (\autoref{table:cp2500a}, \autoref{table:cp2500g}), the advantage available from IVs is even smaller, as we would expect.
For maxed out Pokémon, A\% ranges from 6\%--11\%, increasing as we go down the table.
It is otherwise generally less than 12\%.
These numbers are larger using $\frac{\mathit{BS}}{3}$ than $\sqrt[3]{\BP\,}$; if the latter is used,
  the maximum advantages are contained within 12\%--15\%, and everyone
  else is below 3\%.
\input{out/cp1500a}
\section[\CP2500 bound sorted by $\frac{\mathit{BS}}{3}$ (Ultra League)]%
{$\mathbfit{C\kern -1ptP}$\,2500 bound sorted by $\frac{\mathbfit{BS}}{\mathbf3}$ (Ultra League)}
See \autoref{sec:cp1500a} for important notes.
\input{out/cp2500a}
\section[\CP1500 bound sorted by \protect{$\sqrt[3]{\BP\,}$} (Great League)]%
{$\mathbfit{C\kern -1ptP}$\,1500 bound sorted by $\sqrt[3]{\mathbfit{BP}\;}$ (Great League)\label{sec:cp1500g}}
See \autoref{sec:cp1500a} for important notes.
This table uses the geometric mean $\sqrt[3]{\BP\,}$ instead of the arithmetic mean.
Since Shadow forms don't change under the geometric mean, they have been elided.
\input{out/cp1500g}
\section[\CP2500 bound sorted by \protect{$\sqrt[3]{\BP\,}$} (Ultra League)]%
{$\mathbfit{C\kern -1ptP}$\,2500 bound sorted by $\sqrt[3]{\mathbfit{BP}\;}$ (Ultra League)\label{sec:cp2500g}}
See \autoref{sec:cp1500g} for important notes.
This table uses the geometric mean $\sqrt[3]{\BP{}\,}$ instead of the arithmetic mean.
Since Shadow forms don't change under the geometric mean, they have been elided.
\input{out/cp2500g}
