% oneside ought not be used for printed, bound material!
\documentclass[ebook,10pt,openany,oneside]{memoir}
\title{pgo}
\author{nick black}
%\special{pdf:minorversion 7}
%\usepackage{unicode-math}
\usepackage{amsmath}
\usepackage{fontspec}
\defaultfontfeatures{Scale=MatchUppercase,Ligatures=TeX,Renderer=HarfBuzz}
\usepackage{microtype}
\usepackage{bookmark}
\usepackage{ninecolors}
\usepackage[svgnames,HSB,table]{xcolor}
\selectcolormodel{rgb}
\usepackage[labelfont=bf]{caption} % for \captionsetup
\usepackage{polyglossia}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{tabularray}
\usepackage[export]{adjustbox}
\usepackage{longtable}
\usepackage{tcolorbox}
\tcbuselibrary{most}
\usepackage{listings}
\usepackage{makecell}
\usepackage{varwidth}
\usepackage{ifthen}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{biblatex}
\usepackage[siunitx]{circuitikz}
\usepackage[mode=text]{siunitx}
\setsansfont{Kanit}
\setmainfont{Gentium Book}
\setdefaultlanguage{american}
\newfontfamily\japanesefont[Script=CJK]{Noto Serif CJK JP}
\usepackage{newunicodechar}
\usepackage{realscripts} % typographical enhancement for sub/superscripts
\usepackage{luatex85} % otherwise pdfx blows up =\
\usepackage[a-3u]{pdfx}
\newfontfamily{\symbolfont}{Symbola}
\newunicodechar{🟉}{{\symbolfont🟉}}
\newcolumntype{g}{>{\columncolor{Gray!25}}c} % for alternate column coloring in tabels
\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}} % centering within p-type column

% we're either 6x9" or (less desirably) 8.5x11"
\settrims{0pt}{0pt}
\setstocksize{8.75in}{5.625in}
\settrimmedsize{8.5in}{5.5in}{*}
\setlrmarginsandblock{.75in}{*}{1}
\setulmarginsandblock{1in}{1in}{1}       % 2" of total vertical margin
\setlength{\headwidth}{\textwidth}       % don't go to the end of paperback
\settypeblocksize{7in}{\textwidth}{*}    % want 7in of our 9
\checkandfixthelayout                     % print and check layout % W: Command terminated with space. (1)

\makeatletter
\newcommand*{\textoverline}[1]{$\overline{\hbox{#1}}\m@th$}
\makeatother

\chapterstyle{crosshead}

\renewcommand{\insertchapterspace}{}

\newcommand\Midrule[0]{\midrule}

\definecolor{Shadow}{HTML}{C247FB}
\definecolor{Bug}{HTML}{A6B91A}
\definecolor{Dark}{HTML}{575063}
\definecolor{Dragon}{HTML}{6F35FC}
\definecolor{Electric}{HTML}{F7D02C}
\definecolor{Fairy}{HTML}{D685AD}
\definecolor{Fighting}{HTML}{C22E28}
\definecolor{Fire}{HTML}{EE8130}
\definecolor{Flying}{HTML}{8D87DB}
\definecolor{Ghost}{HTML}{5066A3}
\definecolor{Grass}{HTML}{7AC74C}
\definecolor{Ground}{HTML}{D97A4B}
\definecolor{Ice}{HTML}{96D9D6}
\definecolor{Normal}{HTML}{9CA5AB}
\definecolor{Poison}{HTML}{A33EA1}
\definecolor{Psychic}{HTML}{F95587}
\definecolor{Rock}{HTML}{B6A136}
\definecolor{Steel}{HTML}{5E91A4}
\definecolor{Water}{HTML}{6390F0}


\def\LOGO {
  \textbf{
    \Huge Pokémon GO\\
    \Large A Quantitative Approach\\
  }}
\frontmatter
\begin{document}
  \pagestyle{empty}
  \begin{center}
  \LOGO
  \vspace{0.5in}
    \center{\Large{nick black}}
  \end{center}
  \vfill\hrule
  \begin{center}\textsf{Gold \& Appel Publishing}\end{center}
  \hrule
  \clearpage
  %\input{aux/copyright}
  %\clearpage
  \setcounter{page}{1}
  \pagestyle{plain} % want (roman) page numbers

\clearpage
\vfill
\noindent{}The electronic form of this book is distributed freely.\\
You may make copies, print it out, etc.\\
\\
Should you find it useful, feel encouraged to send an amount you think reasonable to
  \href{mailto:nickblack@linux.com}{nickblack@linux.com} using
  e.g.\ \href{https://paypal.me/dankamongmen}{paypal},
  or to buy the printed edition.
The printed edition does not contain this notice.\\
\\
Thanks! {\textbf{{\symbolfont♥} nick}}
\clearpage
\ifdefined\epub
\else
  \hypertarget{toc}{}%
  \bookmark[dest=toc]{Contents}
  \tableofcontents*
  %\clearpage
  %\hypertarget{lot}{}%
  %\bookmark[dest=lot]{list of tables}
  %\listoftables*
  %\hypertarget{lof}{}%
  %\bookmark[dest=lof]{list of figures} % if we use 3, it changes to i instead of iii [shrug]
  %\listoffigures*
  \fi
\input{foreward}
%--------------------------main text------------------------------------------%
%\pagestyle{companion}
%\input{story}
\input{trainer}
\input{friends}
\input{spawn}
\input{types}
\input{attacks}
\input{species}
\input{pokémon}
\input{max}
\input{damage}
\input{battle}
\input{unbounded}
\input{bounded}
\input{example}
\input{strategy}
\input{simul}
\appendix
\chapter{Optimal League configurations}
\label{chap:optimal}
Forms\footnote{Forms disallowed in League play are not listed.} are sorted by $\sqrt[3]{BP}$, the harmonic mean of $Eff_A$, $Eff_D$, and MHP\@.
This ought not be interpreted to imply that a higher $\sqrt[3]{BP}$ ought
  defeat a lower one; other issues, such as type effectiveness and attack
  choice, are usually more important.
It is possible for multiple sets of IVs to score equally due to the discrete nature
  of MHP\@.
In this case, both optimal sets are listed (there are never more than two).

$A\%$ is the optimal configuration's advantage in terms of $\sqrt[3]{BP}$
  over the pessimal ``reasonable'' configuration, i.e. some IV
  and the highest level attainable given the CP bound.
For CP1500 (\autoref{table:cp1500}), when a form can max out, i.e. is not restricted by the CP bound,
  the maximum advantage tends to be 12\%--15\%, with weaker Pokémon usually
  seeing larger advantages.
It otherwise quickly drops to 4\% or less.
In CP2500 (\autoref{table:cp2500}), the advantage available from IVs is even smaller, as we would expect.
For maxed out Pokémon, A\% ranges from 6\%--11\%, increasing as we go down the table.
It is otherwise generally less than 3\%.
\textbf{FIXME: this is all invalidated by move to arithmetic mean!}

\autoref{table:cp1500}
\textbf{FIXME: add shadow forms}
\input{out/cp1500}
\input{out/cp2500}
\chapter{Species by typings}
\label{chap:speciesbytype}
Forms are sorted by typing, and within typing by Pokédex number.
I provide the base stats and their geometric mean, evolutionary line,
 Pokédex number, an image, cost to purify and to teach a second charged
 move, walk distance (\autoref{subsec:getcandy}), catch and flee rates,
 region, generation, optimal IV+level configurations for CP1500-
 and CP2500-bounded play, weather and Mega boost, and whether the
 form is available as Shiny and/or Shadow.
Forms have been combined when distinguished only visually.

\textbf{FIXME: need index by name}

\textbf{infobox for each species still needs: attacks, genus, evolves from, evolves to,
           requirements to evolve, catch rate, flee rate, region, generation, shadow?, dmax?,
           cost to purify/teach, walk distance, weather boost, candy/items to evolve,
           whether mega etc. forms exist}
\input{out/species}
\chapter{Users of attacks}
\label{chap:attackemployers}
It is often useful to build a team around attacks, rather than Pokémon.
This appendix lists the Pokémon which can learn and use a given attack.
Pokémon in italics will \textit{not} have STAB with the attack.

\textbf{FIXME: need index attacks by name}

\section{Fast attacks}
\label{sec:usersfast}
Fast attacks are sorted by type, then duration, then by the product of DPT and EPT\@.
\input{out/fastusers}

\section{Charged attacks}
\label{sec:userscharged}
Charged attacks are sorted by type, then DPE\@.
\input{out/chargedusers}

\chapter{Permanent evolutions}
\label{chap:allevolutions}
\input{out/evolutions}

\backmatter
\input{about}
\end{document}
