\chapter{Species}
Pokémon are partitioned into species.
Each species is identified by its own Pokédex number.
As we will see, this number is the only thing guaranteed to be shared
 among all the population of a species.

\section{Forms}
Some species boast multiple forms and genders, all sharing one common Pokédex entry.
Form is usually merely a visual distinction, but it is sometimes reflected
 in stats or even typing.
All Pokémon of a given form share the same Attack (ATK), Defense (DEF), and
 Stamina (STA) base statistics.
These statistics are always positive integers.

Common forms include:
\begin{itemize}
  \item Mega, a temporary form with greater ATK and DEF (see \autoref{section:mega}).
  \item Dynamax and Gigantamax, temporary forms with the ability to use
          powerful attacks (see \autoref{section:dmax} and \autoref{section:gmax}).
  \item Shiny, a visually distinctive form (see \autoref{section:shiny}).
  \item The regional forms Alolan, Galarian, Hisuian, and Paldean.
\end{itemize}
Gender, if differentiated at all, is usually a small change to
 appearance or call, but it (like other forms) also affects a few evolutionary
 paths.

\section{Evolution}
Some species can, under the correct conditions, evolve into others.
We call the set of Pokémon related by evolution operations a genus.
Change of form is not an evolution, since the species remains the same.
We call a Pokémon that has evolved $N$ times a Stage $N+1$ Pokémon.
No Pokémon of Stage 4 or greater currently exist.
Evolution does not necessarily preserve typing, i.e. typing is not always constant within genera.

\textbf{FIXME: table of heterogeneous genera}

\textbf{FIXME: item-assisted evolutions}

\section{Shadow and Purified Pokémon}
Team Rocket uses ``Shadow'' Pokémon modified for more attack
 and less defense capability than their base forms (for more details,
 see \autoref{chap:damage}).
When captured from Team Rocket, they enter their own Shadow Pokédex.
Shadow Pokémon cost more to power up at the same Level (120\% the Stardust).
Captured Shadow Pokémon always know the Charged Attack Frustration.
Only during certain special events can this Charged Attack be replaced,
 at which time a Charged TM is required as normal.
It's a pretty terrible Charged Attack, and this really degrades the
 Shadow Pokémon until it can be replaced.
A Shadow Pokémon can be taught a second Charged Attack at any
 time, though this is also more expensive than for the normal form.
Frustration is preserved across Evolution.

\section{Species by type}
\textbf{FIXME: graphs of stat distribution by type}

\input{out/species}
