\chapter{Trainers}
A Pokémon GO account is synonymous with its associated Trainer.

New Trainers start at Level 1 (of 50) with 0 XP\@.
A new Level is conferred as soon as the Trainer's XP meets or exceeds
  the Level's XP threshold (see \autoref{table:xp40}).
\begin{table}[ht]
\begin{center}
\begin{tabular}{r r r|r r r}
  Level & kXP & ΔkXP & Level & kXP & ΔkXP \\
\Midrule
1 & 0 & 0 & 21 & 260 & 50 \\
2 & 1 & 1 & 22 & 335 & 75 \\
3 & 3 & 2 & 23 & 435 & 100 \\
4 & 6 & 3 & 24 & 560 & 125 \\
5 & 10 & 4 & 25 & 710 & 150 \\
6 & 15 & 5 & 26 & 900 & 190 \\
7 & 21 & 6 & 27 & 200 & 1,100 \\
8 & 28 & 7 & 28 & 250 & 1,350 \\
9 & 36 & 8 & 29 & 300 & 1,650 \\
10 & 45 & 9 & 30 & 350 & 2,000 \\
11 & 55 & 10 & 31 & 500 & 2,500 \\
12 & 65 & 10 & 32 & 500 & 3,000 \\
13 & 75 & 10 & 33 & 750 & 3,750 \\
14 & 85 & 10 & 34 & 1,000 & 4,750 \\
15 & 100 & 15 & 35 & 1,250 & 6,000 \\
16 & 120 & 20 & 36 & 7,500 & 7,500 \\
17 & 140 & 20 & 37 & 2,000 & 9,500 \\
18 & 160 & 20 & 38 & 2,500 & 12,000 \\
19 & 185 & 25 & 39 & 3,000 & 15,000 \\
20 & 210 & 25 & 40 & 5,000 & 20,000 \\
\end{tabular}
\caption{Requirements for Trainer Levels 1--40}
\label{table:xp40}
\end{center}
\end{table}
It is possible to advance multiple Levels via a single reward of XP\@.
A Trainer cannot control Pokémon of a Level exceeding the Trainer's by more than 10,
  i.e.\ a Level 22 Trainer cannot power a Pokémon past Level 32.
If such a Pokémon is captured or otherwise received, its Level will be
  reduced to the Trainer's plus 10.
The maximum Pokémon Level is 50, so all Pokémon Levels are available upon
  reaching Trainer Level 40.
Trainer Levels 41 and above have extra requirements beyond XP (see \autoref{table:xp41plus}).

Reaching a new Level is rewarded with items (\autoref{table:levelitems}),
  and sometimes opens up new elements of gameplay (\autoref{table:levelgates}).
Levels 1, 7, 8, 15, and 20 each come with Special Research.

\begin{table}[ht]
\begin{center}
\begin{tabular}{r r r|r r r}
\end{tabular}
\caption{Rewards for achieving Trainer Levels}
\label{table:levelitems}
\end{center}
\end{table}

\begin{table}[ht]
\begin{center}
  \begin{tabular}{r p{0.75\textwidth}}
  Level & Features unlocked \\
\Midrule
  2 & Nenab berries \\
  5 & Potions, Revives, Team selection, appraisal, Gyms, Raids \\
  8 & Razz berries, Team Rocket \\
  10 & Super Potions, Golden Razz berries, Evolution items, Trades, Trainer Battles \\
  12 & Great Balls \\
  15 & Hyper Potions, Fast TMs, Parties \\
  18 & Pinap berries \\
  20 & Silver Pinap berries, Ultra Balls, Mega Evolution \\
  25 & Max Potions, Charged TMs \\
  30 & Max Revives, Route creation \\
  31 & XL Candy \\
  37 & Pokéstop creation \\
\end{tabular}
\caption{Features unlocked by Trainer Levels}
\label{table:levelgates}
\end{center}
\end{table}

Upon reaching Level 5, a Trainer can join one of three Teams: Mystic, Valor, or Instinct.
A Trainer can change Team later by purchasing and making use of a Team Medallion.
The cost is 1,000 Pokécoins, and 365 days must pass between Medallion purchases.

\begin{table}[ht]
\begin{center}
  \begin{tabular}{r r r p{0.65\textwidth}}
Level & MXP & ΔMXP & Special requirements \\
\Midrule
41 & 26 & 6 & 20 Legendary/Mythical power ups,
                      30 Raid wins,
                      200 Pokémon caught in 24 hours,
                      5 Gold medals\\
42 & 33.5 & 7.5 & Evolve all 8 Eevee forms,
                      15 item-assisted evolutions,
                      3 Excellent throws,
                      200 Pokémon caught with berries \\
43 & 42.5 & 9 & 100 Stardust earned,
                      200 effective Charged Attacks,
                      5 Legendary/Mythical catches,
                      5 Platinum medals \\
44 & 53.5 & 11 & 30 Great League wins,
                       30 Ultra League wins,
                       30 Master League wins,
                       20 GO Battle League battles \\
45 & 66.5 & 13 & 100 Rocket Grunts wins,
                       100 Pokémon purified,
                       50 Rocket Leaders wins,
                       10 Platinum medals\\
46 & 82 & 15.5 & 100 Field Research tasks,
                       7 consecutive days w/snapshot,
                       50 Excellent throws,
                       30 Eggs hatched\\
47 & 100 & 18 & 30 Raid wins with heterogenous teams,
                        3{\symbolfont ✯}+ Raid win with CP1500-bounded team,
                        3 power ups to Max CP,
                        20 Platinum medals\\
48 & 121 & 21 & 10 Souvenirs from buddy,
                        300 hearts with buddy,
                        200km walked with buddy,
                        8 weeks with 25km walked\\
49 & 146 & 25 & 10 trades of 300km+ catch distance,
                        50 Lucky Pokémon received in trades,
                        500 Gifts sent,
                        35 Platinum medals\\
50 & 176 & 30 & 999 Excellent throws,
                        5 consecutive Legendary/Mythical catches,
                        3 Rocket Leader wins with CP2500-bounded teams,
                        Rank 10 in GO Battle League\\
\end{tabular}
\caption{Requirements for Trainer Levels 41--50}
\label{table:xp41plus}
\end{center}
\end{table}

All game mechanics, including the ability to fully power up Pokémon,
 are available upon reaching Level 40.
Levels 43, 45, 48, and 50 provide extra Challenge quests.
Levels 43, 45, 47, 49, and 50 open some cosmetic options for Trainer avatars.

\section{Capacities}
A Trainer enters this world with no Pokémon, no Pokécoins, no Stardust,
  no Max Particles, no Candy of any kind, no Mega Energy of any kind,
  no Primal Energy of any kind, and no Fusion Energy of any kind.
The Trainer has carrying capacity sufficient for 300 Pokémon; additional
  capacity can be added in increments of 50 up to a total of 10,500.
Each upgrade costs 200 Pokécoins.
A full upgrade will thus currently run you 40,800 Pokécoins.

The Trainer has a Bag capable of storing 350 items (strangely, not all items
  count against this limit, including Gifts, Stickers, and the Postcard Book).
The Bag comes with an Infinite Incubator, a Camera, 2 Incense, and 50 Poké Balls,
  and thus space for 296 new items.
The Bag can be upgraded to store up to 8,800 items.
Here again, each upgrade costs 200 Pokécoins, and increases capacity by 50 items.
The maximum Bag is thus 33,800 Pokécoins.

The Trainer has a Postcard Book with seven pages, each capable of retaining
 50 Postcards, for an inital limit of 350.
Pages can be added for 100 Pokécoins each, up to a maximum of 40 Pages
 supporting 2,000 total Postcards.
Such a monster will cost 3,300 Pokécoins.

The Trainer can carry up to 20 Gifts at a time. and up to 25 of each sticker.
To date, no one has found any use for the stickers.

The Trainer can accumulate no more than 9999 units of a given Mega Energy,
 Primal Energy, or Fusion Energy.

No limit is known for Candy nor Candy XL.

\section{Collecting Pokémon}
During the initial tour with Professor Willow, you are able to catch one of
 Bulbasaur, Charmander, Squirtle, or Pikachu.
You will be provided an infinite supply of Pokéballs for this encounter, and
 the catch rate (see \autoref{chap:spawn}) is set to 100\%.

\textbf{FIXME}

\section{Collecting XP}
\textbf{FIXME}

\section{Collecting Stardust}
\textbf{FIXME}

\section{Collecting Candy}
A Pokémon can be ``transferred to the Professor'' from your collection,
 whereupon its soul is converted into a tasty Candy (certain special
 Pokémon, including Kubfu, Marshadow and Zygarde, cannot be harvested).
Whenever a Pokémon is captured, Candy is awarded. The amount depends on
 the evolutionary state of the captured Pokémon, and the type of Berry
 applied at the time of capture.
\textbf{FIXME}

\section{Collecting Mega and Primal Energy}
The relevant Energies can be acquired by winning a Mega Raid of that type,
  walking with a Mega Evolved or Primal Reverted Buddy, and completing
  certain research tasks.

\section{Medals}
There is a large set of Medals, each of which can be achieved at the Bronze,
 Silver, Gold, or Platinum level.
\textbf{FIXME}
